---
title: "Mini-Project #04"
subtitle: "Just the Fact(-Check)s, Ma’am!" 
code-fold: true 
toc: true 
toc-depth: 10
execute: 
  echo: true 
  warning: false 
  message: false
---


## Data Acquisition


```{r results='hide'}

library(tidyverse)
library(httr2)
library(rvest)
library(lubridate)
library(DT)
library(scales)
library(infer)



Dir <- file.path("data", "mp04")
if (!dir.exists(Dir)) {
  dir.create(Dir, showWarnings = FALSE, recursive = TRUE)
}


CES <- file.path(Dir, "ces_levels.rds")
CESR <- file.path(Dir, "ces_revisions.rds")

func_ces <- function() {
  
  
  if (file.exists(CES)) {
    return(readRDS(CES))
  }
  
  
  response <- request("https://data.bls.gov/pdq/SurveyOutputServlet") |>
    req_method("POST") |>
    req_headers(
      "User-Agent" = "Mozilla/5.0 (Windows NT 10.0; Win64; x64) AppleWebKit/537.36 (KHTML, like Gecko) Chrome/120.0.0.0 Safari/537.36",
      "Content-Type" = "application/x-www-form-urlencoded"
    ) |>
    req_body_form(
      request_action = "get_data",
      reformat = "true",
      from_results_page = "true",
      from_year = "1979",
      to_year = "2025",
      Go.x = "13",
      Go.y = "10",
      initial_request = "false",
      data_tool = "surveymost",
      series_id = "CES0000000001",
      original_annualAveragesRequested = "false"
    ) |>
    req_timeout(120) |>
    req_perform()
  
  
  html_content <- resp_body_html(response)
  tables <- html_content |> 
    html_elements("table") |> 
    html_table()
  
  
  ces2 <- tables[[2]] |>
    pivot_longer(
      cols = Jan:Dec,
      names_to = "month",
      values_to = "level"
    ) |>
    mutate(
      level = as.numeric(str_remove_all(level, ",")),
      date = ym(paste(Year, month))
    ) |>
    select(date, level) |>
    drop_na() |>
    arrange(date)
  
  saveRDS(ces2, CES)
  message("Saved ", nrow(ces2), " observations to cache")
  
  return(ces2)
}


func_cesr <- function() {
  
 
  if (file.exists(CESR)) {
    return(readRDS(CESR))
  }
  
  message("Fetching CES revisions from BLS website...")
  
  response <- request("https://www.bls.gov/web/empsit/cesnaicsrev.htm") |>
    req_headers(
      "User-Agent" = "Mozilla/5.0 (Windows NT 10.0; Win64; x64) AppleWebKit/537.36 (KHTML, like Gecko) Chrome/120.0.0.0 Safari/537.36"
    ) |>
    req_timeout(60) |>
    req_perform()
  
  
  html_content <- resp_body_html(response)
  
  years <- 1979:2025
  
  cesr2 <- years |>
    map(function(yr) {
      extract_year_revisions(html_content, yr)
    }) |>
    list_rbind()
  
  
  saveRDS(cesr2, CESR)
  message("Saved ", nrow(cesr2), " revision observations to cache")
  
  return(cesr2)
}

CES <- func_ces() 


CESR <- func_cesr() 

if ("date" %in% names(CES)) {
  CES <- CES |> rename(Date = date, Level = level)
}
if ("date" %in% names(CESR)) {
  CESR <- CESR |> rename(Date = date, Original = original, Final = final, Revision = revision)}


```


## Task 1: Download CES Total Nonfarm Payroll

```{r}

datatable(
  CES,
  caption = 'CES Total Nonfarm Payroll Employment (Seasonally Adjusted)',
  colnames = c('Date', 'Level'),
  rownames = FALSE,
  options = list(
    pageLength = 10,
    scrollY = '500px',
    scrollCollapse = TRUE,
    searching = FALSE
  )) 
  



```

## Task 2: Download CES Revisions Tables



```{r}
datatable(
  CESR,
  caption = 'CES Employment Revisions',
  colnames = c('Date', 'Original', 'Final', 'Revision'),
  rownames = FALSE,
  options = list(
    pageLength = 10,
    scrollY = '500px',
    scrollCollapse = TRUE,
    searching = FALSE
  )) 
```

## Task 3: Data Exploration and Visualization
### Table Join


```{r}

CES_combined <- CES |>
  left_join(CESR, by = "Date") |>
  arrange(desc(Date))

```


### Statistics


```{r}


mean_level <- CES_combined |>
  specify(response = Level) |>
  calculate(stat = "mean")

median_level <- CES_combined |>
  specify(response = Level) |>
  calculate(stat = "median")


sd_level <- CES_combined |>
  specify(response = Level) |>
  calculate(stat = "sd")

summary_stats <- data.frame(
  Statistic = c("Mean", "Median", "Standard Deviation"),
  Value = c(mean_level$stat, median_level$stat, sd_level$stat)
)

datatable(
  summary_stats,
  caption = 'Summary Statistics for CES Employment Level',
  rownames = FALSE,
  options = list(
    dom = 't',
    pageLength = 3
  )
) |>
  formatRound(columns = 'Value', digits = 2)

CES_with_metrics <- CES_combined %>%
  mutate(
    Positive_Revision = Revision > 0,
    Negative_Revision = Revision < 0,
    Revision_Pct = (Revision / Original) * 100,
    Absolute_Revision = abs(Revision_Pct) > 50
  )

revision_summary <- data.frame(
  Metric = c(
    "Proportion of Positive Revisions",
    "Proportion of Negative Revisions",
    "Proportion of Large Revisions (>50%)"
  ),
  Value = c(
    mean(CES_with_metrics$Positive_Revision, na.rm = TRUE),
    mean(CES_with_metrics$Negative_Revision, na.rm = TRUE),
    mean(CES_with_metrics$Absolute_Revision, na.rm = TRUE)
  )
)

datatable(
  revision_summary,
  caption = 'Summary Statistics for Revision Metrics',
  rownames = FALSE,
  options = list(
    dom = 't',
    pageLength = 5
  )
) |>
  formatPercentage(columns = 'Value', digits = 2, 
                   rows = c(1, 2)) |>
  formatRound(columns = 'Value', digits = 2)
              





```

### Visualization

```{r}

ggplot(CES, aes(x = Date, y = Level)) +
  geom_line(color = "steelblue") +
  labs(
    title = "CES Total Nonfarm Payroll Employment",
    subtitle = "Seasonally Adjusted",
    x = "Date",
    y = "Employment Level"
  ) +
  theme_bw()


```

```{r}




ggplot(CESR, aes(x = Date, y = Revision)) +
  geom_col(aes(fill = Revision > 0)) +
  scale_fill_manual(values = c("red", "steelblue"), 
                    labels = c("Downward", "Upward")) +
  labs(
    title = "CES Employment Revisions",
    x = "Date",
    y = "Revision (thousands)",
    fill = "Direction"
  ) +
  theme_minimal() +
  geom_hline(yintercept = 0, linetype = "dashed")


```

###  Statistical Inference
There is no statistically significant difference in employment revisions between the pre-2020 period and post-2020 period. The p-value is .487 which is greater than .05. The upper and lower bounds cross zero. We cannot conclude that there is a systemic difference.  

```{r}

CES_with_metrics <- CES_combined |>
  mutate(
    Year = as.numeric(format(Date, "%Y")),
    Post2020 = Year >= 2020
  )

revision_ttest <- CES_with_metrics |>
  t_test(response = Revision,
         explanatory = Post2020,
         order = c(TRUE, FALSE),
         alternative = "two-sided")


datatable(
  revision_ttest,
  caption = 'T-Test Results: Revisions Pre-2020 vs Post-2020',
  rownames = FALSE,
  options = list(
    dom = 't',
    pageLength = 1
  )
) |>
  formatRound(columns = c('statistic', 't_df', 'p_value', 'estimate', 
                          'lower_ci', 'upper_ci'), 
              digits = 3)



```

```{r}







```

